\vspace{-2em} \textbf{ Mapping stellar surfaces\\ II: An interpretable Gaussian process model for light curves }

### 2021-01-26 paper2/ms.tex
misc updates, finished paper 2

@@

\begin{abstract}
    The use of Gaussian processes (GPs) in stellar light curve modeling has recently
    become almost ubiquitous, given their ease of use and flexibility.
    GPs excel in particular at marginalization
    over the stellar signal in cases where the
    variability due to starspots rotating in and out of view is treated
    as a nuisance, such as in exoplanet transit modeling.
    However, in cases where the starspot signal is of primary interest,
    GPs have thus far been less useful: it is often difficult to
    relate their hyperparameters to the physical properties of interest, such
    as the size, contrast, and latitudinal distribution of the spots.
    Instead, it is common practice to explicitly model the effect
    of individual starspots on the light curve and attempt to infer their
    properties via optimization or posterior inference. Unfortunately,
    this process is
    degenerate, ill-posed, and often computationally intractable when
    applied to stars with more than a few spots and/or to ensembles of many
    light curves.
    %
    In this paper, we derive a closed-form expression for the
    mean and covariance of a Gaussian process model that describes
    the light curve of a rotating, evolving stellar surface
    conditioned on a given distribution of
    star spot sizes, contrasts, and latitudes.
    %
    We show that our GP is unbiased,
    allowing one to correctly infer physical parameters of interest from one
    or more stellar light curves, including
    the typical radii and the mean and variance of the latitude
    distribution of starspots.
    %
    Our GP has far-ranging implications for understanding the variability and
    magnetic activity of stars from both light curves and radial velocity (RV)
    measurements, as well as for robustly modeling correlated noise in both transiting
    and RV exoplanet searches.
    %
    Our implementation is efficient, user-friendly, and open source, available
    as the \Python package \starryprocess.
    \href{https://github.com/rodluger/starry_process}{\color{linkcolor}\faGithub}
\end{abstract}

@@

When the goal is to understand
the stellar surface itself, however, GPs have been less useful to date.
While there have been several recent papers

\xxx{...}
the exponential sine squared kernel \citep[Equation~\ref{eq:expsinsq}][]{Wang2012},
quasi-periodic kernel \citep{Angus2018},
or the \celerite \textsf{RotationTerm} \citep{ForemanMackey2017}.
quasi-periodic cosine kernel\citep[e.g.,][]{Perger2020}
\xxx{...}

Instead, the common approach is that discussed in the previous section:
to forward model the stellar surface and its effect on the light curve, then
seek to constrain the surface properties by solving the inverse problem.
As we discussed above, however, a unique solution to the surface map of
the star from a rotational light curve is not possible without the use of
aggressive (and often \emph{ad hoc}) priors. The degeneracies at play
make it effectively impossible for one to know the exact configuration
of starspots and other features on the surface of a star from its rotational
light curve alone.

However, it is hardly ever the case that this is actually our end goal.
After all, physics can be used to predict properties of stellar surfaces at a fairly
high level: i.e., typical spot sizes, active spot latitudes, or approximate
timescales on which spots evolve
\citep[e.g.,][]{Schuessler1996,Solanki2006,Cantiello2019}.
We are hardly ever interested in the
\emph{particular} properties of a \emph{particular} spot, as we wouldn't really
know what to do with that information! Instead, we often treat
(whether explicitly or not)
the properties of a star spot as a draw from some parent distribution
controlling (say) the average and spread in the radii of the spots.
The parameters controling this distribution are the ones that we can
predict with physics; they are therefore also the ones we are usually
interested in.

@@

\textbf{In this paper, we derive an exact, closed-form expression for
    the Gaussian approximation to the marginal likelihood of a light curve
    conditioned on the statistical properties of starspots.} Our Gaussian
process analytically marginalizes over the (degenerate, often unknowable)
distributions of properties of individual starspots, revealing the
constraints imposed on the bulk spot properties without the need to
explicitly model or sample over properties of individual spots. It inherits
the speed, ease-of-use, and all other properties of traditionally-used
GPs, with the added benefit of direct physical interpretability of its
hyperparameters.
%

The paper is organized as follows:
we present an overview of the derivation of the GP in \S\ref{sec:gp}
and a suite of tests on synthetic data to show the model is calibrated
in \S\ref{sec:calibration}. We discuss our GP from an information
theory perspective in \S\ref{sec:discussion}, focusing on
the various degeneracies our model can and cannot break when applied
to ensembles of light curves. In \S\ref{sec:implementation} we present
our implementation of the algorithm in the open-source \starryprocess package
alongside speed and stability tests, and in \S\ref{sec:conclusions} we
summarize our results. Most of the math behind the algorithm
is presented in the Appendix, followed by a series of supplementary figures
(discussed in \S\ref{sec:calibration}). Appendix~\ref{sec:notation} discusses the
notation we adopt throughout the paper and Table~\ref{tab:variables}
lists the main symbols and variables, with links to their definitions.

Finally, we note that all of the figures in this paper were auto-generated
using the Azure Pipelines continuous integration (CI) service, which
ensures they are up to date with the latest version of the
\starryprocess code. In particular, icons next to each of the figures \codeicon \,
link to the exact script used to generate them to ensure the reproducibility
of our results. As in \citet{PaperI}, the principal equations are
accompanied by
``unit tests'': \textsf{pytest}-compatible test scripts associated
with the principal equations that pass (fail) if the equation is correct (wrong),
in which case a clickable \testpassicon \, (\testfailicon) is shown next to the equation
label.
In most cases, the validity of an equation is gauged by comparison to
a numerical solution. Like the figure scripts, the equation unit tests are
run on Azure Pipelines upon every commit of the code.

@@

In this section, we present an overview of the computation of our Gaussian
process, which boils down to computing the mean and covariance of the stellar
flux conditioned on certain physical properties of the star and its star spot
distribution. Most of the math is folded into the Appendix for readability;
readers may want to refer to Appendix~\ref{sec:notation} in particular for
a discussion of the notation and conventions we adopt.

@@

Before we dive in, it is useful to introduce the spherical harmonics,
a set of orthogonal functions on the surface of the sphere which we will use
to describe the intensity field on the surface of a star
(Figure~\ref{fig:ylms}). As we will see below,
the spherical harmonics are a particularly convenient basis in which to
describe star spot distributions%
\footnote{There are, of course, drawbacks to using this basis:
    in particular, the spherical harmonics are smooth, continuous functions that
    struggle (at finite degree $l$) to capture high resolution features such as
    small starspots. We discuss this point at length in \S\ref{sec:tinyspots},
    where we show that our model is useful even when applied to stars with spots
    smaller than the effective resolution of the GP.}, as they will allow us to compute
moments of the intensity distribution analytically. Of more immediate concern,
\citet{Luger2019} showed that there is a linear relationship between the
spherical harmonic expansion of a stellar surface and the total disk-integrated
flux $\mathbf{f}$ (i.e., the light curve)
one would observe as the star rotates about a fixed axis.
If the stellar surface intensity is described by a spherical harmonic
coefficient vector $\mathbf{y}$ (up to a certain degree $l_\mathrm{max}$),
the flux is given by
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}}(I, P, \mathbf{u}) \, \mathbf{y}
    \quad,
\end{align}
%
where $\mathbf{1}$ is the ones vector and
$\pmb{\mathcal{A}}$ is the \starry
design matrix, a purely linear operator that transforms from the spherical
harmonic basis to the flux basis; it is
a function of the stellar inclination $I$, the stellar
rotation period $P$, and the stellar limb darkening coefficients $\mathbf{u}$
(see Appendix~\ref{sec:starry} for details).

@@

\subsection{The baseline problem}

\xxx{BLA BLA BLA}

since we must be careful to distinguish between the
\emph{sample} mean and the \emph{process} mean. When normalizing a light
curve by the mean, the operation we perform is
%
\begin{align}
    \label{eq:ftilde}
    \tilde{\mathbbb{f}\hspace{0.2em}} & = \frac{\mathbbb{f}}{\left<\mathbb{f}\right>}
    \quad,
\end{align}
%
where $\tilde{\mathbbb{f}\hspace{0.2em}}$ is the normalized, unit-mean light curve,
$\mathbbb{f}$ is the measured light curve (in detector counts), and
$\left<\mathbb{f}\right>$ is the \emph{sample} mean: i.e., the average value of
a given star's light curve (which we model as a sample from our GP).
This may be close to but is in general different from the \emph{process} mean,
$\pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)$, since the mean of
a draw from the GP is itself normally distributed with a variance that scales
with the GP variance.%
\footnote{
    Importantly, the sample mean and process mean will be different even in the
    absence of measurement error! In other words, the mean flux of a given
    star (i.e., the sample mean)
    will in general be different from the mean flux \emph{across all stars} with similar
    surface properties (the process mean).
}

Computing the covariance of the normalized process is tricky, especially
because the normalized process is \emph{not} strictly Gaussian: the distribution
has heavy tails due to the fact that $\tilde{\mathbbb{f}\hspace{0.2em}}$ diverges as
the sample mean approaches zero. In fact, because of these tails, the covariance
of the normalized process is formally \emph{infinite}, since the probability of
drawing a sample whose mean is arbitrarily close to zero is finite.

If this is all starting to sound like a bad idea, that's because it is!
A much safer approach is to resist the temptation to normalize the light curve
and instead model the unknown multiplicative amplitude as a latent variable. However,
this would require an extra parameter \emph{for every light curve}, so the
computational savings we achieved by marginalizing out the inclination
would be gone. Fortunately, in practice, the variance of a stellar light curve
is usually small compared to its mean: stellar variability amplitudes are
typically at the level of a few percent or lower. When this is the case,
the probability of drawing a GP sample whose mean is close to zero is
extremely small, and we can make use of the approximate expression derived
in \citet{Luger2021} for the covariance of a normalized Gaussian process:
%
\begin{proof}{test_normgp}
    \label{eq:SigmaTilde}
    \tilde{\pmb{\Sigma}}
    & \approx
    \frac{A}{\mu^2} \pmb{\Sigma} +
    z \Big(
    (A + B) \, (\mathbf{1} - \mathbf{q}) \, (\mathbf{1} - \mathbf{q})^\top
    - A \, \mathbf{q} \, \mathbf{q}^\top
    \Big)
    \quad,
\end{proof}
%
where
%
\begin{align}
    \label{eq:z}
    z & \equiv \frac{\left< \Sigma \right>}{\mu^2}
\end{align}
%
is the ratio of the average element in $\pmb{\Sigma}$
to the square of the mean of the Gaussian process,
$\mathbf{q}$ is the ratio of the average of each row in $\pmb{\Sigma}$
to the average element in $\pmb{\Sigma}$, and $A$, $B$ are
order unity and zero scalars, respectively,
given by the optimally-truncated diverging series
%
\begin{align}
    \label{eq:baseline_alpha}
    A
     & \equiv
    \sum\limits_{i=0}^{i_\mathrm{max}}
    \frac{(2i + 1)!}{2^i \, i!}
    z^i
    \\[1em]
    \label{eq:baseline_beta}
    B
     & \equiv
    \sum\limits_{i=0}^{i_\mathrm{max}}
    \frac{2i(2i + 1)!}{2^i \, i!}
    z^i
    \quad,
\end{align}
%
where $i_\mathrm{max}$ is the largest value for which the series coefficient at $i_\mathrm{max}$ is
smaller than the coefficient at $i_\mathrm{max} - 1$. In the expressions above, it is
assumed that the mean $\pmb{\mu}$ is constant, i.e., $\pmb{\mu} = \mu\, \mathbf{1}$.
Since our Gaussian process is azimuthally isotropic (i.e., no preferred
longitude), that is the case throughout this paper.

\xxx{VERY IMPORTANT for this paper. The variance of the light curves, which scales their
    amplitudes, depends on the process MEAN, which is unobservable. Studies that look at
    light curve ``amplitudes'' are therefore biased, even if only slightly.}

@@

\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/normgp.pdf}
        \oscaption{normgp}{%
            An example of a flux covariance matrix $\pmb{\Sigma}$ for a
            \starry process (left) and the corresponding covariance of the
            normalized process (right), computed from Equation~(\ref{eq:SigmaTilde}).
            In addition to an offset and an overall scaling relative to the
            original covariance matrix, the covariance of the normalized process
            is discernibly non-stationary (see Figure~\ref{fig:nonstationarity}).
            \label{fig:normgp}
        }
    \end{centering}
\end{figure}

@@

We model the mean-normalized flux $\tilde{\mathbbb{f}\hspace{0.2em}}$
(Equation~\ref{eq:ftilde}) as a Gaussian process:
%
\begin{align}
    \tilde{\mathbbb{f}\hspace{0.2em}}
    \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \sim
    \mathcal{N}\left(
    \mathbf{1},
    \tilde{\pmb{\Sigma}} \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \right)
    \quad.
\end{align}
%
The hyperparameters of the GP are
the stellar rotation period $P$,
the vector of limb darkening
coefficients $\mathbf{u}$, and the vector of parameters describing the
spot distribution
%
\begin{align}
    \pmb{\theta}_\bullet
     & =
    \left(
    n
    \,\,\,
    c
    \,\,\,
    \mu_\phi
    \,\,\,
    \sigma_\phi
    \,\,\,
    r
    \right)^\top
    \quad,
\end{align}
%
consisting of the number of spots $n$, their contrast $c$,
the mode $\mu_\phi$ and standard deviation $\sigma_\phi$ of the latitude
distribution, and the radius of the spots $r$.
The quantity $\tilde{\pmb{\Sigma}}$ is the covariance of the (normalized) process
(Equation~\ref{eq:SigmaTilde}), which is a straightforward correction to the
true covariance of the process, which accounts for
changes in scale and phase introduced by the common process of
normalizing light curves to a mean of unity. It depends on the
true (constant) mean $\mu$
and true covariance $\pmb{\Sigma}$,
given by Equations~(\ref{eq:mu_marg}) and (\ref{eq:cov_marg}),
respectively.
Those expressions in turn depend on the inclination expectation integrals
$e_I$ (Appendix~\ref{sec:inc-mom1}) and $\mathbf{E}_I$
(Appendix~\ref{sec:inc-mom2}). Those, in turn, depend on the first and
second moments of the distribution of spherical harmonic coefficient vectors,
$\mathrm{E}\left[ \mathbbb{y} \, \big| \, \pmb{\theta}_\bullet \right]$
and
$\mathrm{E}\left[ \mathbbb{y} \, \mathbbb{y}^\top \, \big| \, \pmb{\theta}_\bullet \right]$,
given by Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively.
To compute those, we must evaluate four nested integrals
(Equations~\ref{eq:e1}--\ref{eq:e4} for the first moment
and \ref{eq:E1}--\ref{eq:E4} for the second moment), corresponding to integrals
over the radius, latitude, longitude, and contrast distributions, respectively.
The computation of these integrals is discussed at length in Appendix~\ref{sec:integrals}.

While lengthy (and quite tedious), all of the computations described above rely
on equations whose solutions have a closed form.%
%
\footnote{The exception to this is the
    normalization correction (\S\ref{sec:baseline}), which depends on a
    fast-to-evaluate series and thus adds negligible overhead to the computation.}
%
Moreover, most of the terms in the expectation vectors and matrices may
be computed recursively, and many may be pre-computed, as they do not
depend on user inputs.
%
It is therefore possible to evaluate $\tilde{\pmb{\Sigma}}$ in an
efficient manner. In \S\ref{sec:implementation} we discuss our
implementation of the algorithm in a user-friendly \Python package.

@@

Figure~\ref{fig:calibration_default_data} shows a synthetic dataset generated
from the default values listed in Table~\ref{tab:synthetic}. While the
light curves correspond to surfaces with the same statistical spot properties,
they all look qualitatively different: the mapping from star spot properties
to flux is nontrivial. In the inference step below, we assume we observe
only these 50 light curves (the figure only shows 49 of them),
with no knowledge of the inclination of any individual
star, and attempt to infer the spot properties.

\subsection{Inference}
\label{sec:calibration-inference}
We use our \Python-based implementation of the GP (\S\ref{sec:implementation})
to perform inference on the synthetic datasets. For simplicity, we assume we
know the value of the period $P$, which is fixed at unity for all stars, as well
as the value of the limb darkening coefficients (fixed at zero for the default
run). In practice these will not be known exactly; we discuss this further in
\S\ref{sec:other-marg}. Since we explicitly marginalize over the inclinations
of individual stars, the only quantities we must constrain are the five
parameters in the spot parameter vector $\pmb{\theta}_\bullet$
(Equation~\ref{eq:thetaspot}). We experimented with three different methods
for doing posterior inference on our synthetic datasets: No-U Turn Sampling,
a variant of Hamiltonian Monte Carlo
\citep[NUTS;][]{Duane1987,Hoffman2011}, automatic differentiation
variational inference \citep[ADVI;][]{Kucukelbir2016,Blei2016}, and nested sampling \citep{Skilling2004,Skilling2006}.
We obtained the best performance using the nested sampling algorithm
implemented in the \textsf{dynesty} package \citep{Speagle2020}, so that is what
we will use below.

% HACK HACK HACK HACK HACK
\pagebreak
% HACK HACK HACK HACK HACK

@@

@@

We conclude our discussion with a list of several notes and caveats that should be kept
in mind when using our algorithm and/or its \Python implementation.

@@

          play. In particular, spots centered on the poles are always in

@@

>>>>

\defcitealias{PaperI}{Paper I}

@@

\begin{abstract}
    The use of Gaussian processes (GPs) in stellar light curve modeling has recently
    become almost ubiquitous, given their ease of use and flexibility.
    GPs excel in particular at marginalization
    over the stellar signal in cases where the
    variability due to starspots rotating in and out of view is treated
    as a nuisance, such as in exoplanet transit modeling.
    However, in cases where the starspot signal is of primary interest,
    GPs have thus far been less useful: it is often difficult to
    relate their hyperparameters to the physical properties of interest, such
    as the size, contrast, and latitudinal distribution of the spots.
    Instead, it is common practice to explicitly model the effect
    of individual starspots on the light curve and attempt to infer their
    properties via optimization or posterior inference. Unfortunately,
    this process is
    degenerate, ill-posed, and often computationally intractable when
    applied to stars with more than a few spots and/or to ensembles of many
    light curves.
    %
    In this paper, we derive a closed-form expression for the
    mean and covariance of a Gaussian process model that describes
    the light curve of a rotating, evolving stellar surface
    conditioned on a given distribution of
    starspot sizes, contrasts, and latitudes.
    %
    We show that our GP is unbiased,
    allowing one to correctly infer physical parameters of interest from one
    or more stellar light curves, including
    the typical radii and the mean and variance of the latitude
    distribution of starspots.
    %
    Our GP has far-ranging implications for understanding the variability and
    magnetic activity of stars from both light curves and radial velocity (RV)
    measurements, as well as for robustly modeling correlated noise in both transiting
    and RV exoplanet searches.
    %
    Our implementation is efficient, user-friendly, and open source, available
    as the \Python package \starryprocess.
    \href{https://github.com/rodluger/starry_process}{\color{linkcolor}\faGithub}
\end{abstract}

@@

When the goal is to understand
the stellar surface itself, however, GPs have been less useful to date.
While recent papers have proposed several custom
GP kernels that excel at modeling rotational variability, such as
the exponential sine squared kernel \citep[Equation~\ref{eq:expsinsq};][]{Wang2012},
the quasi-periodic kernel \citep{Angus2018}, and
the \celerite \textsf{RotationTerm} \citep{ForemanMackey2017},
the connection between the kernel hyperparameters and the physical spot
properties is often qualitative. Recently, \citet{Perger2020} proposed
an empirically-motivated quasi-periodic cosine kernel that can be used
to infer spot lifetimes in the absence of differential rotation. However,
an explicit connection between the GP hyperparameters and the size and
spatial distribution of spots has not been explored in the literature.

Instead, a common approach to inferring starspot properties from
photometric datasets is
to forward model the stellar surface and its effect on the light curve, then
seek to constrain the surface properties by solving the inverse problem.
We discussed this approach in \citet{PaperI} \citepalias[hereafter][]{PaperI},
where we argued a unique solution to the surface map of
the star from a rotational light curve is not possible without the use of
aggressive (and often \emph{ad hoc}) priors. The degeneracies at play
make it effectively impossible for one to know the exact configuration
of starspots and other features on the surface of a star from its rotational
light curve alone.

However, it is hardly ever the case that this is actually our end goal.
After all, physics can be used to predict properties of stellar surfaces at a fairly
high level: i.e., typical spot sizes, active spot latitudes, or approximate
timescales on which spots evolve
\citep[e.g.,][]{Schuessler1996,Solanki2006,Cantiello2019}.
We are hardly ever interested in the
\emph{particular} properties of a \emph{particular} spot, as we wouldn't really
know what to do with that information! Instead, we often treat
(whether explicitly or not)
the properties of a starspot as a draw from some parent distribution
controlling (say) the average and spread in the radii of the spots.
The parameters controling this distribution are the ones that we can
predict with physics; they are therefore also the ones we are usually
interested in.

@@

\textbf{In this paper, we derive an exact, closed-form expression for
    the Gaussian approximation to the marginal likelihood of a light curve
    conditioned on the statistical properties of starspots.} Our Gaussian
process analytically marginalizes over the degenerate and often unknowable
distributions of properties of individual starspots, revealing the
constraints imposed on the bulk spot properties without the need to
explicitly model or sample over properties of individual spots. It inherits
the speed, ease-of-use, and all other properties of traditionally-used
GPs, with the added benefit of direct physical interpretability of its
hyperparameters.
%
In particular, \textbf{our GP is aimed at ensemble analyses of many light curves.}
As we showed in \citetalias{PaperI}, the joint information content of
the light curves of many stars can break many of the degeneracies in the
mapping problem. We encourage readers to read \citetalias{PaperI} to
better understand this and other points regarding the information theory
behind stellar rotational light curves.

The present paper is organized as follows:
we present an overview of the derivation of the GP in \S\ref{sec:main}
and a suite of tests on synthetic data to show the model is calibrated
in \S\ref{sec:calibration}. We discuss our results and the limitations
of our model in \S\ref{sec:discussion} and present straightforward
extensions of the GP, including its application to time-variable
surfaces, in \S\ref{sec:extensions}.
In \S\ref{sec:conclusions} we summarize our results and discuss
topics we will address in future papers in this series.

Most of the math behind the algorithm
is presented in the Appendix, followed by a series of supplementary figures
(discussed in \S\ref{sec:calibration}). Appendix~\ref{sec:notation} discusses the
notation we adopt throughout the paper and Table~\ref{tab:variables}
lists the main symbols and variables, with links to their definitions.
The algorithm developed in this paper is fully implemented in the \starryprocess code, which is
available on \href{https://github.com/rodluger/starry_process}{GitHub}
and is described in more detail in \citet{JOSSPaper}.
%

Finally, we note that all of the figures in this paper were auto-generated
using the Azure Pipelines continuous integration (CI) service, which
ensures they are up to date with the latest version of the
\starryprocess code. In particular, icons next to each of the figures \codeicon \,
link to the exact script used to generate them to ensure the reproducibility
of our results. As in \citetalias{PaperI}, the principal equations are
accompanied by
``unit tests'': \textsf{pytest}-compatible test scripts associated
with the principal equations that pass (fail) if the equation is correct (wrong),
in which case a clickable \testpassicon \, (\testfailicon) is shown next to the equation
label.
In most cases, the validity of an equation is gauged by comparison to
a numerical solution. Like the figure scripts, the equation unit tests are
run on Azure Pipelines upon every commit of the code.

@@

In this section, we present an overview of the computation of our Gaussian
process, which boils down to computing the mean and covariance of the stellar
flux conditioned on certain physical properties of the star and its starspot
distribution. Most of the math is folded into the Appendix for readability;
readers may want to refer to Appendix~\ref{sec:notation} in particular for
a discussion of the notation and conventions we adopt.

@@

Before we dive in, it is useful to introduce the spherical harmonics,
a set of orthogonal functions on the surface of the sphere which we will use
to describe the intensity field on the surface of a star
(Figure~\ref{fig:ylms}). As we will see below,
the spherical harmonics are a particularly convenient basis in which to
describe starspot distributions%
\footnote{There are, of course, drawbacks to using this basis:
    in particular, the spherical harmonics are smooth, continuous functions that
    struggle (at finite degree $l$) to capture high resolution features such as
    small starspots. We discuss this point at length in \S\ref{sec:tinyspots},
    where we show that our model is useful even when applied to stars with spots
    smaller than the effective resolution of the GP.}, as they will allow us to compute
moments of the intensity distribution analytically. Of more immediate concern,
\citet{Luger2019} showed that there is a linear relationship between the
spherical harmonic expansion of a stellar surface and the total disk-integrated
flux $\mathbf{f}$ (i.e., the light curve)
one would observe as the star rotates about a fixed axis.
If the stellar surface intensity is described by a spherical harmonic
coefficient vector $\mathbf{y}$ (up to a certain degree $l_\mathrm{max}$),
the flux is given by
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}}(I, P, \mathbf{u}) \, \mathbf{y}
    \quad,
\end{align}
%
where $\mathbf{1}$ is the ones vector and
$\pmb{\mathcal{A}}$ is the \starry
design matrix, a purely linear operator that transforms from the spherical
harmonic basis to the flux basis; it is
a function of the stellar inclination $I$, the stellar
rotation period $P$, and the stellar limb darkening coefficients $\mathbf{u}$
(see Appendix~\ref{sec:starry} for details).

@@

\subsection{Normalization correction}
\label{sec:baseline}

In \citetalias{PaperI} we discussed a subtle but important point
about stellar light curves: the common procedure of normalizing
light curves to their mean or median level changes the covariance
structure of the data, since it correlates all the observations
in a nontrivial way.
%
When normalizing a light
curve by the mean,%
\footnote{In practice, the expressions derived here also work well
    for median-normalized light curves, since the distribution of the GP sample median
    is usually close to the distribution of the sample mean.}
the operation we perform is
%
\begin{align}
    \label{eq:ftilde}
    \tilde{\mathbbb{f}\hspace{0.2em}} & = \frac{\mathbbb{f}}{\left<\mathbb{f}\right>}
    \quad,
\end{align}
%
where $\tilde{\mathbbb{f}\hspace{0.2em}}$ is the normalized, unit-mean light curve,
$\mathbbb{f}$ is the measured light curve (in detector counts), and
$\left<\mathbb{f}\right>$ is the \emph{sample} mean: i.e., the average value of
a given star's light curve (which we model as a sample from our GP).
This may be close to but is in general different from the \emph{process} mean,
$\pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)$, since the mean of
a draw from the GP is itself normally distributed with a variance that scales
with the GP variance.%
\footnote{
    Importantly, the sample mean and process mean will be different even in the
    absence of measurement error! In other words, the mean flux of a given
    star (i.e., the sample mean)
    will in general be different from the mean flux \emph{across all stars} with similar
    surface properties (the process mean).
}

When modeling normalized light curves, we must correct our expression for
the covariance matrix $\pmb{\Sigma}$ of the GP.
Computing the new covariance matrix is tricky, especially
because the normalized process is \emph{not} strictly Gaussian: the distribution
of normalized light curves
has heavy tails due to the fact that $\tilde{\mathbbb{f}\hspace{0.2em}}$ diverges as
the sample mean approaches zero. In fact, because of these tails, the covariance
of the normalized process is formally \emph{infinite}, since the probability of
drawing a sample whose mean is arbitrarily close to zero is finite.

If this is all starting to sound like a bad idea, that's because it is!
A much safer approach is to resist the temptation to normalize the light curve
and instead model the (unknown) amplitude of the data as a multiplicative
latent variable. However,
this would require an extra parameter \emph{for every light curve}, so the
computational savings we achieved by marginalizing out the inclination
would be gone. Fortunately, in practice, the variance of a stellar light curve
is usually small compared to its mean: stellar variability amplitudes are
typically at the level of a few percent or lower. When this is the case,
the probability of drawing a GP sample whose mean is close to zero is
extremely small, and we can make use of the approximate expression derived
in \citet{Luger2021} for the covariance of a normalized Gaussian process:
%
\begin{proof}{test_normgp}
    \label{eq:SigmaTilde}
    \tilde{\pmb{\Sigma}}
    & \approx
    \frac{A}{\mu^2} \pmb{\Sigma} +
    z \Big(
    (A + B) \, (\mathbf{1} - \mathbf{q}) \, (\mathbf{1} - \mathbf{q})^\top
    - A \, \mathbf{q} \, \mathbf{q}^\top
    \Big)
    \quad,
\end{proof}
%
where
%
\begin{align}
    \label{eq:z}
    z & \equiv \frac{\left< \Sigma \right>}{\mu^2}
\end{align}
%
is the ratio of the average element in $\pmb{\Sigma}$
to the square of the mean of the Gaussian process,
$\mathbf{q}$ is the ratio of the average of each row in $\pmb{\Sigma}$
to the average element in $\pmb{\Sigma}$, and $A$, $B$ are
order unity and zero scalars, respectively,
given by the optimally-truncated diverging series
%
\begin{align}
    \label{eq:baseline_alpha}
    A
     & \equiv
    \sum\limits_{i=0}^{i_\mathrm{max}}
    \frac{(2i + 1)!}{2^i \, i!}
    z^i
    \\[1em]
    \label{eq:baseline_beta}
    B
     & \equiv
    \sum\limits_{i=0}^{i_\mathrm{max}}
    \frac{2i(2i + 1)!}{2^i \, i!}
    z^i
    \quad,
\end{align}
%
where $i_\mathrm{max}$ is the largest value for which the series coefficient at $i_\mathrm{max}$ is
smaller than the coefficient at $i_\mathrm{max} - 1$. In the expressions above, it is
assumed that the mean $\pmb{\mu}$ is constant, i.e., $\pmb{\mu} = \mu\, \mathbf{1}$.
Since our Gaussian process is azimuthally isotropic (i.e., no preferred
longitude), that is the case throughout this paper.

% The variance of the light curves, which scales their
% amplitudes, depends on the process MEAN, which is unobservable. Studies that look at
% light curve ``amplitudes'' are therefore biased, even if only slightly.

@@

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/normgp.pdf}
        \oscaption{normgp}{%
            An example of a flux covariance matrix $\pmb{\Sigma}$ for a
            \starry process (left) and the corresponding covariance of the
            normalized process (right), computed from Equation~(\ref{eq:SigmaTilde}).
            In addition to an offset and an overall scaling relative to the
            original covariance matrix, the covariance of the normalized process
            is discernibly non-stationary.
            \label{fig:normgp}
        }
    \end{centering}
\end{figure}

@@

We model the mean-normalized flux $\tilde{\mathbbb{f}\hspace{0.2em}}$
(Equation~\ref{eq:ftilde}) as a Gaussian process:
%
\begin{align}
    \tilde{\mathbbb{f}\hspace{0.2em}}
    \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \sim
    \mathcal{N}\left(
    \mathbf{1},
    \tilde{\pmb{\Sigma}} \left(P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \right)
    \quad.
\end{align}
%
The hyperparameters of the GP are
the stellar rotation period $P$,
the vector of limb darkening
coefficients $\mathbf{u}$, and the vector of parameters describing the
spot distribution
%
\begin{align}
    \pmb{\theta}_\bullet
     & =
    \left(
    n
    \,\,\,
    c
    \,\,\,
    \mu_\phi
    \,\,\,
    \sigma_\phi
    \,\,\,
    r
    \right)^\top
    \quad,
\end{align}
%
consisting of the number of spots $n$, their contrast $c$,
the mode $\mu_\phi$ and standard deviation $\sigma_\phi$ of the latitude
distribution, and the radius of the spots $r$.
The quantity $\tilde{\pmb{\Sigma}}$ is the covariance of the normalized process
(Equation~\ref{eq:SigmaTilde}), which is a straightforward correction to the
true covariance of the process, accounting for
changes in scale and phase introduced by the common process of
normalizing light curves to a mean of unity. It depends on the
true (constant) mean $\mu$
and true covariance $\pmb{\Sigma}$,
given by Equations~(\ref{eq:mu_marg}) and (\ref{eq:cov_marg}),
respectively.
Those expressions in turn depend on the inclination expectation integrals
$e_I$ (Appendix~\ref{sec:inc-mom1}) and $\mathbf{E}_I$
(Appendix~\ref{sec:inc-mom2}). Those, in turn, depend on the first and
second moments of the distribution of spherical harmonic coefficient vectors,
$\mathrm{E}\left[ \mathbbb{y} \, \big| \, \pmb{\theta}_\bullet \right]$
and
$\mathrm{E}\left[ \mathbbb{y} \, \mathbbb{y}^\top \, \big| \, \pmb{\theta}_\bullet \right]$,
given by Equations~(\ref{eq:exp_y_sep}) and (\ref{eq:exp_yy_sep}), respectively.
To compute those, we must evaluate four nested integrals
(Equations~\ref{eq:e1}--\ref{eq:e4} for the first moment
and \ref{eq:E1}--\ref{eq:E4} for the second moment), corresponding to integrals
over the radius, latitude, longitude, and contrast distributions, respectively.
The computation of these integrals is discussed at length in Appendix~\ref{sec:integrals}.

While lengthy (and quite tedious), all of the computations described above rely
on equations whose solutions have a closed form.%
%
\footnote{The exception to this is the
    normalization correction (\S\ref{sec:baseline}), which depends on a
    fast-to-evaluate series and thus adds negligible overhead to the computation.}
%
Moreover, most of the terms in the expectation vectors and matrices may
be computed recursively, and many may be pre-computed, as they do not
depend on user inputs.
%
It is therefore possible to evaluate $\tilde{\pmb{\Sigma}}$ in an
efficient manner. In the companion paper \citep{JOSSPaper}, we discuss our
implementation of the algorithm in a user-friendly \Python package.

@@

Figure~\ref{fig:calibration_default_data} shows a synthetic dataset generated
from the default values listed in Table~\ref{tab:synthetic}. While the
light curves correspond to surfaces with the same statistical spot properties,
they all look qualitatively different: the mapping from starspot properties
to flux is nontrivial. In the inference step below, we assume we observe
only these 50 light curves (the figure only shows 49 of them),
with no knowledge of the inclination of any individual
star, and attempt to infer the spot properties.

% HACK HACK HACK HACK HACK
\vfill
\pagebreak
% HACK HACK HACK HACK HACK

\subsection{Inference}
\label{sec:calibration-inference}
We use our \Python-based implementation of the GP \citep{JOSSPaper}
to perform inference on the synthetic datasets. For simplicity, we assume we
know the value of the period $P$, which is fixed at unity for all stars, as well
as the value of the limb darkening coefficients (fixed at zero for the default
run). In practice these will not be known exactly; we discuss this further in
\S\ref{sec:other-marg}. Since we explicitly marginalize over the inclinations
of individual stars, the only quantities we must constrain are the five
parameters in the spot parameter vector $\pmb{\theta}_\bullet$
(Equation~\ref{eq:thetaspot}). We experimented with three different methods
for doing posterior inference on our synthetic datasets: No-U Turn Sampling,
a variant of Hamiltonian Monte Carlo
\citep[NUTS;][]{Duane1987,Hoffman2011}, automatic differentiation
variational inference \citep[ADVI;][]{Kucukelbir2016,Blei2016}, and nested sampling \citep{Skilling2004,Skilling2006}.
We obtained the best performance using the nested sampling algorithm
implemented in the \textsf{dynesty} package \citep{Speagle2020}, so that is what
we will use below.

@@

\subsubsection{Synthetic likelihoods, random fields, and approximate inference}

@@

We conclude our discussion with a list of several notes and caveats that should be kept
in mind when using our algorithm and its \Python implementation.

@@

          play \citepalias{PaperI}. In particular, spots centered on the poles are always in

@@

    \item \textbf{Typically, an ensemble of at least $M \sim 50$ light curves
